\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-14 smurf_paper.tex
add support for latex+eps images instead of pdflatex+pdf images
First run convert_images to change all of the .pdf files into .eps.

Then uncomment the relevant block at the start of smurf_paper.tex. This
macro '\imagefile{image}' just tacks a .eps or .pdf on the end.
graphicx already knows how to deal with .eps vs .pdf suffix in image files
initial submission to MN
I put everything, including the information to fill out the on-line forms,
in the submission_mnras subdirectory. Currently smurf_paper.tex is set to
use postscript instead of PDF to follow their submission guidelines.

%% Recommended by MN for handling hyperlinks
\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}

% this is needed to get pdftex to work properly!
\usepackage[pdftex]{graphicx}

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area on several
different time scales, and at different scan angles); and (ii) minimal
accelerations to reduce turn-around overheads (which could be quite
large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
larger than the array footprint a rectangular ``PONG'' pattern is
used, in which the boresight travels in approximately straight lines
and ``bounces'' off the edges at 45\,degree angles until the area is
uniformly filled in. It is also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking. For smaller areas (of order the
array footprint, or pont sources), in which the PONG turn-around
overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
used.  Here the telescope moves in a circle, whose centre also slowly
traces out a small circle. For a more complete description of the
\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals over-plotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{pspec_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{pspec.pdf}
\caption{Bolometer power spectral densities for the same two data sets
  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
  noise slightly and clarify some features. Four of the most sensitive
  bolometers have been selected at each wavelength, and the
  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
  as coloured dotted lines (the blue signals are for the same time
  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
  lines are the PSDs of the common-mode signals at each wavelength,
  and the solid coloured lines show the PSDs of the bolometers once
  the common-mode is removed.  Finally, the dashed black lines show
  the spectral shape produced by a point source given the scan speeds
  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
  reference, the top horizontal axes shows the conversion from
  frequency to angular scale. Horizontal dotted lines at $2 \times
  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
  respectively, are provided as a visual reference for the white-noise
  levels. In addition to $1/f$ and white noise components, and line
  features, all of the PSDs exhibit the gradual roll-off of the
  anti-aliasing filter just below the Nyquist frequency.  Left: for
  the S2SRO data there are broad line features in the PSDs at both
  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
  450 and 850\,\micron. Common-mode subtraction removes most of the
  correlated fridge oscillation signal, lowering the $1/f$ knees to
  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
  respectively. Right: data from the fully-commissioned instrument
  tend to have lower $1/f$ knees, and fewer line features. The
  remaining low-frequency noise, now that the fridge oscillations have
  been removed, is considerably more independent from bolometer to
  bolometer than in the S2SRO data (note the larger spread in the
  dotted lines at low frequencies compared to the left-hand plot,
  particularly at 450\,\micron). Finally, this example shows that the
  white noise performance is in fact similar for the two subarrays
  (s4a and s8b) that were used both before and after the upgrades.}
\label{fig:pspec}
\end{figure*}

@@

Raw data from the fully-commissioned instrument (right-hand panel of
Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
compared with the S2SRO data, and it is considerably less correlated
amongst detectors (larger spread in the dotted coloured lines,
particularly noticeable in these 450\,\micron\ data), leading to a
less drastic improvement upon common-mode removal. However, since the
data are less dominated by low-frequency noise to begin with, the
signals generally require less agressive high-pass filtering to
produce maps, and therefore retain larger-scale structures than with
the S2SRO data. Furthermore, there are generally fewer line features
in the PSDs of bolometers in the fully-commissioned
instrument. Finally, these two data sets illustrate that the white
noise performance (NEFD) is in fact similar before and after the
upgrades for these two subarrays (s4a at 450\,micron, and s8b at
850\,\micron). The main improvements are a reduction in correlated and
line noise features mentioned above, the larger number of working
bolometers and field-of-view (approximately a factor of 4).

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{magpickup.pdf}
\caption{Evidence for significant magnetic field pickup for
  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
  panel shows two un-flatfielded (but mean-subtracted and step
  corrected) bolometer time-series from the same column, with a 200
  sample boxcar smooth (approximately 1\,s), illustrating that they
  are dominated by a similar signal with opposite signs. The second
  panel shows the dark squid signal for the column, also
  mean-subtracted and with the same boxcar smooth. The bottom panel
  shows the azimuthal and elevation offsets from the map centre (mean
  azimuth and elevation 171.9$^\circ$ and 68.0$^\circ$,
  respectively). Only the azimuthal signal is obviously correlated
  with the dark squids and bolometer signals, which suggests a
  magnetic field stationary with respect to the telescope dome as the
  source, since its direction with respect to the cryostat only
  changes with azimuthal motion.}
\label{fig:magpickup}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pca.pdf}
\caption{The first twelve components from a principal component
  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
  subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
  data, for the same observation as that used in the right-hand panels
  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
  the top plot shows the time-series of the normalised eigenvectors,
  the middle plot its PSD, and the bottom coloured panels the
  eigenvalues for the bolometers at each wavelength (the amplitudes of
  the eigenvectors in the time-series). For reference, both the mean,
  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
  the bolometers in each subarray are also shown. Most of the
  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
  appears to be dominated by scan-synchronous noise (compare with the
  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
  Components~3, 4, 7 and 11 are transient glitches that only appear in
  a small subset of the bolometers. Component~5 also appears in only a
  few bolometers, and is dominated by a narrow line at 57\,Hz which
  suggests that it is related to the 60\,Hz mains. Components~6 and
  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
  probably aliased high-frequency noise), with some low-frequency
  drifts. Finally, the erratic appearance of the eigenvector for
  Component~12 is typical of a poorly-biased detector.}
\label{fig:pca}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{dimm.pdf}
\caption{Typical map-making algorithm. Raw data (stored in multiple
  files) are read and concatenated into continuous time-series, and
  flatfieded. Based on the scan speed of the telescope, the data are
  down-sampled to match the selected output map pixel size (usually 2
  and 4\,arcsec at 450 and 850\,\micron, respectively, to ensure
  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs DC
  steps and spikes, and fills gaps to ensure continuity. Finally, the
  mean of each bolometer time-series is removed (higher-order
  polynomials may also be used). The iterative portion then begins
  (dashed box): estimating and removing the common-mode signal (the
  combined \model{COM} and \model{GAI} models); applying the
  extinction correction (\model{EXT}); high-pass filtering to remove
  residual independent low-frequency noise (\model{FLT}); estimating
  the map by re-gridding the data, and then removing its projection
  from the time-series (\model{AST}); and finally measuring the noise
  in the residual time-series (\model{NOI}). If the solution has
  converged, the map is written to disk. Otherwise any multiplicative
  factors that may have been applied to the data are inverted (i.e.,
  the extinction correction, \model{EXT}), and then the models for the
  low-frequency noise; the common-mode (\model{COM,GAI}) and high-pass
  filter (\model{FLT}). Each model is then re-estimated in turn until
  \model{AST}, at which point the previous estimate of the
  astronomical signal is added back into the data prior to its
  re-calculation.}
\label{fig:dimm}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{steps1.pdf}
\caption{Examples of steps in time-series data. Steps occur with a wide
range of heights from the large steps shown on the left to the small
steps shown on the right. Large steps are often followed by a brief
``over-shoot'' as shown on the left. In both plots, the black curve is
the uncorrected time-series, and the red curve is the corrected time
series. Samples close to a step are omitted in the corrected time-series.
In the right hand plot, the blue curve is the uncorrected time-series for
a nearby bolometer. The similarity between the red and blue curves shows
that the step correction is performing well.
}
\label{fig:steps1}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad2.pdf}
\caption{The black curve shows a bolometer time-series, padded
with zeros. The red curve shows the time-series after apodisation.}
\label{fig:pad2}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad1.pdf}
\caption{The black curve shows the same bolometer time-series as in
Fig.~\ref{fig:pad2}, again padded with zeros. The red curve shows the
time-series after padding with artificial data.}
\label{fig:pad1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com.pdf}
\caption{A selection of typical bolometer time-series after
flatfielding and removal of a constant baseline. It can be seen that
most bolometers exhibit a common time variation overlaid with other
features. Empty squares indicate the locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pointmaps.pdf}
\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
  (black). (a) Reduction in which only common-mode subtraction is used
  to suppress low-frequency noise, and the reduction is forced to use
  2 iterations (after the first iteration an estimate of the source
  flux is removed, and the noise is measured in the residuals to
  obtain appropriate weighting for the second and final
  iteration). Circular streaks are caused by independent low-frequency
  noise that is not removed by the common-mode (Uranus peak
  0.15\,pW). (b) Same as (a) but using 50 iterations, illustrating the
  degeneracy between large-scale structure and the common-mode (the
  footprint of working bolometers is also shown for reference, Uranus
  peak 0.15\,pW). (c) Reduction in which high-pass filtering above
  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec, as
  indicated) is applied after common-mode removal, but before the map
  estimate. The map-based convergence test is activated and the
  solution halts after 10 iterations, but leaving large-scale ringing
  due to the filter (Uranus peak 0.27\,pW). (d) Same as (c), but now
  the region of the map beyond the white dot-dashed circle is
  constrained to a value of zero until all but the final iteration.
  The map is now extremely flat, there is little attenuation of the
  source flux, and the diffraction pattern is clearly seen (Uranus
  peak 0.27\,pW).}
\label{fig:pointmaps}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{degeneracy.pdf}
\caption{Demonstration of the degeneracy between large-scale
  structures in the map and common-mode removal, corresponding to
  panels (a) and (b) in Fig.~\ref{fig:pointmaps}. The black and grey
  lines in the top panel show the residual signal for a single
  bolometer after 2 and 100 iterations, respectively (they are nearly
  identical; the black line lies beneath the grey line). The green and
  red lines show the difference between the \model{COM} and
  \model{AST} (the signal produced by the current map estimate for a
  given bolometer) model components for that bolometer between
  iterations 2 and 100, respectively. This shows that a strong signal
  has grown over time, and it has equal but opposite signs in the two
  model components, so that they cancel one another. The bottom panel
  shows the scan pattern of the telescope over the same period;
  clearly the \model{COM}/\model{AST} model degeneracy is correlated
  with the elevation offset, and referring to
  Fig.~\ref{fig:pointmaps}, panel (b), this corresponds to the strong
  vertical gradient that has appeared}.
\label{fig:degeneracy}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{lockman_maps.pdf}
\caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
  output of SMURF using high-pass filtering as a pre-processing step,
  followed by 4 iterations using only the \model{COM} model to remove
  residual low-frequency noise. White contours correspond to estimated
  noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
  centre of the map. (b) The whitened map using a filter based on the
  angular power spectrum of the jackknife map. (c) The whitened map
  cross-correlated with the whitened PSF to identify point sources
  [restricted to a lower-noise region, within the area denoted by the
  second contour of panel (a)]. The image shows the \snr\, with
  3.8-$\sigma$ peaks indicated by blue circles (radius 8\,arcsec). The
  orange ``$+$'' signs show the locations of 1.4\,GHz radio sources
  from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
  peaks, 9 are within this search radius of at least one radio
  source. (d) Jackknife map produced from the difference of two maps,
  using the even and odd scan numbers, respectively. Provided that all
  noise sources are statistically uncorrelated between the two halves
  of the data, the map is a plausible realisation of the noise without
  contamination from astronomical sources. (e) The jackknife map
  whitened using the same filter as that in panel (b). The whitened
  jackknife map cross-correlated with the same whitened PSF as in
  panel (c). Again, orange ``$+$'' and blue circles indicate radio
  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
  there is only a single (apparently) significant peak, and it is not
  in the vicinity of a radio source.}
\label{fig:lockman_maps}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_psf.pdf}
\caption{Slice through the angular response to an ideal Gaussian point
  source (solid line) along the R.A. axis following map-making using
  the default blank-field processing with SMURF (dotted line), and
  upon application of the whitening filter (dashed line). The ``map
  filtered'' response is produced by adding the ideal Gaussian to the
  real data near the centre of the Lockman Hole map, and re-reducing
  the data. The resulting dotted line gives the expected shape of a
  source in panel (a) from Fig.~\ref{fig:lockman_maps}. Applying the
  whitening filter (reciprocal of the solid orange line in
  Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
  line (dashed) , which is the expected shape of a source in panel (b)
  from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
  whitened map with this whitened PSF is an effective ``matched
  filter'' for identifying point-like sources, and this smoothed map
  is shown in panel (c) of Fig.~\ref{fig:lockman_maps}.}
\label{fig:lockman_psf}
\end{figure}

@@

%SMURF is highly configurable, making it easy to vary the order
%and types of models that are fit to the data. It is also easy to add
%new models to the existing infrastructure.

>>>>

% --- Uncomment following block for use with pdflatex and pdf images ----------
%% Recommended by MN for handling hyperlinks

%\usepackage[pdftex,pdfpagemode={UseOutlines},bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}

%\usepackage[pdftex]{graphicx}

% -----------------------------------------------------------------------------

% --- Uncomment following block for latex and postscript images ---------------
%% Recommended by MN for handling hyperlinks - PS version

\usepackage[bookmarks,bookmarksopen,colorlinks,linkcolor={blue},citecolor={green},urlcolor={red}]{hyperref}

% -----------------------------------------------------------------------------

@@

In order to reduce the impact of low-frequency noise on the final
maps, \scuba\ scan strategies have been designed to provide: (i) good
cross-linking (visiting every point of the mapped area on several
different time scales, and at different scan angles); and (ii) minimal
accelerations to reduce turn-around overheads (which could be quite
large given the 600\,arcsec\,s$^{-1}$ maximum scan speed). For areas
larger than the array footprint a rectangular ``PONG'' pattern is
used, in which the boresight travels in approximately straight lines
and ``bounces'' off the edges at 45\,degree angles until the area is
uniformly filled in. It is also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking. For smaller areas (of order the
array footprint, or point sources), in which the PONG turn-around
overheads would be large, a ``constant velocity daisy'' (CV Daisy) is
used.  Here the telescope moves in a circle, whose centre also slowly
traces out a small circle. For a more complete description of the
\scuba\ observing modes, see Section~5 in \citet{holland2012} and also
\citet{2010SPIE.7740E..66K}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro}
\includegraphics[width=0.49\linewidth]{bolos_point_mix}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals over-plotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.49\linewidth]{pspec_s2sro}
\includegraphics[width=0.49\linewidth]{pspec}
\caption{Bolometer power spectral densities for the same two data sets
  (before and after upgrades) used in Fig.~\ref{fig:bolos_mix}. The
  PSDs have been boxcar smoothed with a width of 0.1\,Hz to reduce the
  noise slightly and clarify some features. Four of the most sensitive
  bolometers have been selected at each wavelength, and the
  flat-fielded and step-corrected (but otherwise raw) PSDs are shown
  as coloured dotted lines (the blue signals are for the same time
  series as those shown in Fig.~\ref{fig:bolos_mix}).  The solid black
  lines are the PSDs of the common-mode signals at each wavelength,
  and the solid coloured lines show the PSDs of the bolometers once
  the common-mode is removed.  Finally, the dashed black lines show
  the spectral shape produced by a point source given the scan speeds
  for the two observations (120\,arcsec\,s$^{-1}$ in the left-hand
  plot, and 190\,arcsec\,s$^{-1}$ in the right-hand plot); for
  reference, the top horizontal axes shows the conversion from
  frequency to angular scale. Horizontal dotted lines at $2 \times
  10^5$ and $1 \times 10^5$\,pW\,Hz$^{-1}$ at 450 and 850\,\micron,
  respectively, are provided as a visual reference for the white-noise
  levels. In addition to $1/f$ and white noise components, and line
  features, all of the PSDs exhibit the gradual roll-off of the
  anti-aliasing filter just below the Nyquist frequency.  Left: for
  the S2SRO data there are broad line features in the PSDs at both
  wavelengths above $\sim$35\,Hz.  At lower frequencies, the bolometer
  signals exhibit clear $1/f$ knees at approximately 1 and 2\,Hz at
  450 and 850\,\micron. Common-mode subtraction removes most of the
  correlated fridge oscillation signal, lowering the $1/f$ knees to
  approximately 0.2 and 0.7\,Hz at 450 and 850\,\micron,
  respectively. Right: data from the fully-commissioned instrument
  tend to have lower $1/f$ knees, and fewer line features. The
  remaining low-frequency noise, now that the fridge oscillations have
  been removed, is considerably more independent from bolometer to
  bolometer than in the S2SRO data (note the larger spread in the
  dotted lines at low frequencies compared to the left-hand plot,
  particularly at 450\,\micron). Finally, this example shows that the
  white noise performance is in fact similar for the two subarrays
  (s4a and s8b) that were used both before and after the upgrades.}
\label{fig:pspec}
\end{figure*}

@@

Raw data from the fully-commissioned instrument (right-hand panel of
Fig.~\ref{fig:pspec}) generally have less significant $1/f$ noise as
compared with the S2SRO data, and it is considerably less correlated
amongst detectors (larger spread in the dotted coloured lines,
particularly noticeable in these 450\,\micron\ data), leading to a
less drastic improvement upon common-mode removal. However, since the
data are less dominated by low-frequency noise to begin with, the
signals generally require less aggressive high-pass filtering to
produce maps, and therefore retain larger-scale structures than with
the S2SRO data. Furthermore, there are generally fewer line features
in the PSDs of bolometers in the fully-commissioned
instrument. Finally, these two data sets illustrate that the white
noise performance (NEFD) is in fact similar before and after the
upgrades for these two subarrays (s4a at 450\,micron, and s8b at
850\,\micron). The main improvements are a reduction in correlated and
line noise features mentioned above, the larger number of working
bolometers and field-of-view (approximately a factor of 4).

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{magpickup}
\caption{Evidence for significant magnetic field pickup for
  observation 16 on 20100228 at 450\,\micron\ (s4a subarray).  The top
  panel shows two un-flatfielded (but mean-subtracted and step
  corrected) bolometer time-series from the same column, with a 200
  sample boxcar smooth (approximately 1\,s), illustrating that they
  are dominated by a similar signal with opposite signs. The second
  panel shows the dark squid signal for the column, also
  mean-subtracted and with the same boxcar smooth. The bottom panel
  shows the azimuthal and elevation offsets from the map centre (mean
  azimuth and elevation 171.9$^\circ$ and 68.0$^\circ$,
  respectively). Only the azimuthal signal is obviously correlated
  with the dark squids and bolometer signals, which suggests a
  magnetic field stationary with respect to the telescope dome as the
  source, since its direction with respect to the cryostat only
  changes with azimuthal motion.}
\label{fig:magpickup}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pca}
\caption{The first twelve components from a principal component
  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
  subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
  data, for the same observation as that used in the right-hand panels
  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
  the top plot shows the time-series of the normalised eigenvectors,
  the middle plot its PSD, and the bottom coloured panels the
  eigenvalues for the bolometers at each wavelength (the amplitudes of
  the eigenvectors in the time-series). For reference, both the mean,
  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
  the bolometers in each subarray are also shown. Most of the
  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
  appears to be dominated by scan-synchronous noise (compare with the
  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
  Components~3, 4, 7 and 11 are transient glitches that only appear in
  a small subset of the bolometers. Component~5 also appears in only a
  few bolometers, and is dominated by a narrow line at 57\,Hz which
  suggests that it is related to the 60\,Hz mains. Components~6 and
  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
  probably aliased high-frequency noise), with some low-frequency
  drifts. Finally, the erratic appearance of the eigenvector for
  Component~12 is typical of a poorly-biased detector.}
\label{fig:pca}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{dimm}
\caption{Typical map-making algorithm. Raw data (stored in multiple
  files) are read and concatenated into continuous time-series, and
  flatfieded. Based on the scan speed of the telescope, the data are
  down-sampled to match the selected output map pixel size (usually 2
  and 4\,arcsec at 450 and 850\,\micron, respectively, to ensure
  adequate sampling of the \scuba\ PSFs). A cleaning stage repairs DC
  steps and spikes, and fills gaps to ensure continuity. Finally, the
  mean of each bolometer time-series is removed (higher-order
  polynomials may also be used). The iterative portion then begins
  (dashed box): estimating and removing the common-mode signal (the
  combined \model{COM} and \model{GAI} models); applying the
  extinction correction (\model{EXT}); high-pass filtering to remove
  residual independent low-frequency noise (\model{FLT}); estimating
  the map by re-gridding the data, and then removing its projection
  from the time-series (\model{AST}); and finally measuring the noise
  in the residual time-series (\model{NOI}). If the solution has
  converged, the map is written to disk. Otherwise any multiplicative
  factors that may have been applied to the data are inverted (i.e.,
  the extinction correction, \model{EXT}), and then the models for the
  low-frequency noise; the common-mode (\model{COM,GAI}) and high-pass
  filter (\model{FLT}). Each model is then re-estimated in turn until
  \model{AST}, at which point the previous estimate of the
  astronomical signal is added back into the data prior to its
  re-calculation.}
\label{fig:dimm}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{steps1}
\caption{Examples of steps in time-series data. Steps occur with a wide
range of heights from the large steps shown on the left to the small
steps shown on the right. Large steps are often followed by a brief
``over-shoot'' as shown on the left. In both plots, the black curve is
the uncorrected time-series, and the red curve is the corrected time
series. Samples close to a step are omitted in the corrected time-series.
In the right hand plot, the blue curve is the uncorrected time-series for
a nearby bolometer. The similarity between the red and blue curves shows
that the step correction is performing well.
}
\label{fig:steps1}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad2}
\caption{The black curve shows a bolometer time-series, padded
with zeros. The red curve shows the time-series after apodisation.}
\label{fig:pad2}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{pad1}
\caption{The black curve shows the same bolometer time-series as in
Fig.~\ref{fig:pad2}, again padded with zeros. The red curve shows the
time-series after padding with artificial data.}
\label{fig:pad1}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com}
\caption{A selection of typical bolometer time-series after
flatfielding and removal of a constant baseline. It can be seen that
most bolometers exhibit a common time variation overlaid with other
features. Empty squares indicate the locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pointmaps}
\caption{Multiple reductions of a 450\,\micron\ CV daisy scan of
  Uranus, all scaled between $-$0.002\,pW (white) and +0.002\,pW
  (black). (a) Reduction in which only common-mode subtraction is used
  to suppress low-frequency noise, and the reduction is forced to use
  2 iterations (after the first iteration an estimate of the source
  flux is removed, and the noise is measured in the residuals to
  obtain appropriate weighting for the second and final
  iteration). Circular streaks are caused by independent low-frequency
  noise that is not removed by the common-mode (Uranus peak
  0.15\,pW). (b) Same as (a) but using 50 iterations, illustrating the
  degeneracy between large-scale structure and the common-mode (the
  footprint of working bolometers is also shown for reference, Uranus
  peak 0.15\,pW). (c) Reduction in which high-pass filtering above
  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec, as
  indicated) is applied after common-mode removal, but before the map
  estimate. The map-based convergence test is activated and the
  solution halts after 10 iterations, but leaving large-scale ringing
  due to the filter (Uranus peak 0.27\,pW). (d) Same as (c), but now
  the region of the map beyond the white dot-dashed circle is
  constrained to a value of zero until all but the final iteration.
  The map is now extremely flat, there is little attenuation of the
  source flux, and the diffraction pattern is clearly seen (Uranus
  peak 0.27\,pW).}
\label{fig:pointmaps}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{degeneracy}
\caption{Demonstration of the degeneracy between large-scale
  structures in the map and common-mode removal, corresponding to
  panels (a) and (b) in Fig.~\ref{fig:pointmaps}. The black and grey
  lines in the top panel show the residual signal for a single
  bolometer after 2 and 100 iterations, respectively (they are nearly
  identical; the black line lies beneath the grey line). The green and
  red lines show the difference between the \model{COM} and
  \model{AST} (the signal produced by the current map estimate for a
  given bolometer) model components for that bolometer between
  iterations 2 and 100, respectively. This shows that a strong signal
  has grown over time, and it has equal but opposite signs in the two
  model components, so that they cancel one another. The bottom panel
  shows the scan pattern of the telescope over the same period;
  clearly the \model{COM}/\model{AST} model degeneracy is correlated
  with the elevation offset, and referring to
  Fig.~\ref{fig:pointmaps}, panel (b), this corresponds to the strong
  vertical gradient that has appeared}.
\label{fig:degeneracy}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{lockman_maps}
\caption{Reduction of a blank-field survey: the Lockman Hole. (a) Raw
  output of SMURF using high-pass filtering as a pre-processing step,
  followed by 4 iterations using only the \model{COM} model to remove
  residual low-frequency noise. White contours correspond to estimated
  noise levels of 1.25, 2.5 and 5.0 times the minimum noise at the
  centre of the map. (b) The whitened map using a filter based on the
  angular power spectrum of the jackknife map. (c) The whitened map
  cross-correlated with the whitened PSF to identify point sources
  [restricted to a lower-noise region, within the area denoted by the
  second contour of panel (a)]. The image shows the \snr\, with
  3.8-$\sigma$ peaks indicated by blue circles (radius 8\,arcsec). The
  orange ``$+$'' signs show the locations of 1.4\,GHz radio sources
  from \citet{owen2008} with $S\gsim15$\,$\mu$Jy. Of the 10 submm
  peaks, 9 are within this search radius of at least one radio
  source. (d) Jackknife map produced from the difference of two maps,
  using the even and odd scan numbers, respectively. Provided that all
  noise sources are statistically uncorrelated between the two halves
  of the data, the map is a plausible realisation of the noise without
  contamination from astronomical sources. (e) The jackknife map
  whitened using the same filter as that in panel (b). The whitened
  jackknife map cross-correlated with the same whitened PSF as in
  panel (c). Again, orange ``$+$'' and blue circles indicate radio
  sources and 3.8-$\sigma$ peaks, respectively. Unlike panel (c),
  there is only a single (apparently) significant peak, and it is not
  in the vicinity of a radio source.}
\label{fig:lockman_maps}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_psf}
\caption{Slice through the angular response to an ideal Gaussian point
  source (solid line) along the R.A. axis following map-making using
  the default blank-field processing with SMURF (dotted line), and
  upon application of the whitening filter (dashed line). The ``map
  filtered'' response is produced by adding the ideal Gaussian to the
  real data near the centre of the Lockman Hole map, and re-reducing
  the data. The resulting dotted line gives the expected shape of a
  source in panel (a) from Fig.~\ref{fig:lockman_maps}. Applying the
  whitening filter (reciprocal of the solid orange line in
  Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
  line (dashed) , which is the expected shape of a source in panel (b)
  from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
  whitened map with this whitened PSF is an effective ``matched
  filter'' for identifying point-like sources, and this smoothed map
  is shown in panel (c) of Fig.~\ref{fig:lockman_maps}.}
\label{fig:lockman_psf}
\end{figure}

@@
